---
title: "PHClaims table dependencies"
author: "Alastair Matheson (Assessment, Policy Development, and Evaluation Unit)"
date: "02/24/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require("pacman")) {install.packages("pacman")}
pacman::p_load(tidyverse, glue, DiagrammeR)

#### Set up color schema
color_load_raw <- "SeaGreen1"
color_stage <- "SeaGreen2"
color_final <- "SeaGreen3"
color_archive <- "SeaGreen4"
color_ref <- "Bisque1"
color_metadata <- "Bisque2"
color_file <- "Azure"
color_edge <- "MidnightBlue"
color_edge_qa <- "CornflowerBlue"
color_font_light <- "Snow"
color_font_dark <- "DarkOliveGreen"

#### Pull in and set up data
# input <- data.table::fread("https://raw.githubusercontent.com/PHSKC-APDE/claims_data/main/claims_db/phclaims/table_dependencies.csv")
input <- data.table::fread("C:/code/claims_data/claims_db/phclaims/table_dependencies.csv")

# Modify overall data
input <- input %>%
  mutate(to = paste(schema, table, sep = "."),
         from = case_when(
           !is.na(parent_schema) & parent_schema != "" ~ paste(parent_schema, parent_table, sep = "."),
           !is.na(parent_file) & parent_file != "" ~ parent_file))

# Set up overall nodes
nodes <- input %>%
  distinct(to, schema) %>%
  rename(node = to) %>%
  bind_rows(., distinct(input, from, parent_schema) %>% 
              filter(!is.na(from)) %>%
              rename(node = from, schema = parent_schema))  %>%
  arrange(node) %>%
  mutate(fillcolor = case_when(
           schema == "load_raw" ~ color_load_raw,
           schema == "stage" ~ color_stage,
           schema == "final" ~ color_final,
           schema == "archive" ~ color_archive,
           schema == "ref" ~ color_ref,
           schema == "metadata" ~ color_metadata,
           is.na(schema) | schema == "" ~ color_file,
           TRUE ~ "Black"),
         fontcolor = case_when(
           schema %in% c("final", "archive") ~ color_font_light,
           TRUE ~ color_font_dark),
         node_label = str_replace(node, "\\.", "\\.\\\\n")
         ) %>%
  distinct(node, fillcolor, fontcolor, node_label) %>%
  mutate(id = seq(1, nrow(.))) %>%
  select(id, node, fillcolor, fontcolor, node_label)

# Set up overall edges
edges <- input %>%
  filter(!is.na(to) & !is.na(from)) %>%
  mutate(color = case_when(
           is.na(qa_dependency) ~ color_edge,
           TRUE ~ color_edge_qa),
         penwidth = case_when(
           is.na(qa_dependency) ~ 1,
           TRUE ~ 0.5),
         rel = case_when(
           is.na(qa_dependency) ~ "depends_on",
           TRUE ~ "helps_qa")
         ) %>%
  select(from, to, color, penwidth, rel) %>%
  arrange(from) %>%
  # Keep IDs to set up edges
  left_join(., select(nodes, node, id),
            by = c("from" = "node")) %>%
  rename(from_id = id) %>%
  left_join(., select(nodes, node, id),
            by = c("to" = "node")) %>%
  rename(to_id = id) %>%
  mutate(edge = paste(from_id, to_id, sep = '->'),
         # Need to be factors for DiagrammeR to work
         from = factor(from, levels = nodes$node),
         to = factor(to, levels = nodes$node))


# Set up function to restrict to desired nodes
node_filter <- function(edge_df) {
  output <- edge_df %>%
  mutate(node = as.character(from)) %>%
  bind_rows(., mutate(edge_df, node = as.character(to)) %>% select(node)) %>%
  distinct(node) %>%
  left_join(., nodes, by = "node")
}

# Set up a function to create the Graphviz text
gv_maker <- function(node_df, edge_df) {
  gv_nodes <- glue_collapse(
    glue_data(node_df, "{id} [fillcolor = {fillcolor}, fontcolor = {fontcolor}, label = '{node_label}']"),
    sep = '\n')
  
  gv_edges <- glue_collapse(glue_data(edge_df, "{edge} [color = {color}, pendwidth = {penwidth}]"), sep = '; ')
  
  gv_txt <- glue('
  digraph dependencies {{
  
  ## Graph statement
  graph [overlap = false]
  labelloc = t;
  rankdir = LR;
  
  ## Set up overall formats
  node[shape = rectangle, style = filled]
  
  ## Set up legend
  subgraph cluster_legend {{
  
    label = "Key";
    color = Crimson;
    
    subgraph cluster_col1 {{
      label = "";
      style = invis;
    
      kc1[fillcolor = {color_load_raw}, label = ""];
      k1[shape=plaintext, style=solid, width=2, label = "load_raw schema\\r"]
      k1->kc1[style = invis]
      
      kc2[fillcolor = {color_stage}, label = ""];
      k2[shape=plaintext, style=solid, width=2, label = "stage schema\\r"]
      k2->kc2[style = invis]
      
      kc3[fillcolor = {color_final}, label = ""];
      k3[shape=plaintext, style=solid, width=2, label = "final schema\\r"]
      k3->kc3[style = invis]
      
      kc4[fillcolor = {color_archive}, label = ""];
      k4[shape=plaintext, style=solid, width=2, label = "archive schema\\r"]
      k4->kc4[style = invis]
    }}


    subgraph cluster_col2 {{
      label = "";
      style = invis;
      
      kc5[fillcolor = {color_ref}, label = ""];
      k5[shape=plaintext, style=solid, width=2, label = "ref schema\\r"]
      k5->kc5[style = invis]
      
      kc6[fillcolor = {color_metadata}, label = ""];
      k6[shape=plaintext, style=solid, width=2, label = "metadata schema\\r"]
      k6->kc6[style = invis]
      
      kc7[fillcolor = {color_file}, label = ""];
      k7[shape=plaintext, style=solid, width=2, label = "file source\\r"]
      k7->kc7[style = invis]
    }}
    
    
    subgraph cluster_col3 {{
      label = "";
      style = invis;
      
      kc8[style = invis]
      k8[shape=plaintext, style=solid, width=2, label = "depends on\\r"]
      k8->kc8[color = {color_edge}]
      
      kc9[style = invis]
      k9[shape=plaintext, style=solid, width=2, label = "helps QA\\r"]
      k9->kc9[color = {color_edge_qa}]
    }}
    
    # Add invisible edges so the legend has columns
    kc1->k5[style = invis]
    kc5->k8[style = invis]

  }}
  
  ## Set up main graph
  subgraph cluster_main {{
    # Set up nodes
    {gv_nodes}
  
    # Set up edges
    {gv_edges}
  }}

  
  }}')
  
  return(gv_txt)
}



### NB
# After experimenting with various methods of using DiagrammeR, the best results
# came from using glue to generate Graphviz code rather than the create_graph and 
# render_graph functions. 

```

## Purpose

This is an attempt to document the table dependencies that exist in the PHClaims database. This will allow users to better understand how code/variable changes
will impact downstream tables. Because of the volume of tables, the diagrams are split into sections.

Future plans for this documentation include turning it into an interactive document that would allow users to select a specific table and see all the related tables.

Code to create this document is found here: <https://github.com/PHSKC-APDE/claims_data/blob/main/claims_db/phclaims/table_dependencies.Rmd>


## APCD 
### Claims
```{r apcd_claim, echo=FALSE}
# set up apcd_elig only
apcd_claim_edges <- edges %>% 
  filter(str_detect(from, "apcd_claim") | str_detect(to, "apcd_claim"))

# See which nodes should be included
apcd_claim_nodes <- node_filter(apcd_claim_edges)
  
# Set up Graphviz text
apcd_claim_gv <- gv_maker(node_df = apcd_claim_nodes, edge_df = apcd_claim_edges)

# Make graph
if (length(apcd_claim_gv)) {
  grViz(apcd_claim_gv)
} else {
  message("No tables in this category")
}
```

### Eligibility
```{r apcd_elig, echo=FALSE}
# set up apcd_elig only
apcd_elig_edges <- edges %>% 
  filter(str_detect(from, "apcd_elig") | str_detect(to, "apcd_elig"))

# See which nodes should be included
apcd_elig_nodes <- node_filter(apcd_elig_edges)
  
# Set up Graphviz text
apcd_elig_gv <- gv_maker(node_df = apcd_elig_nodes, edge_df = apcd_elig_edges)

# Make graph
if (length(apcd_elig_gv)) {
  grViz(apcd_elig_gv)
} else {
  message("No tables in this category")
}
```


## Medicaid 
### Claims
```{r mcaid_claim, echo=FALSE}
# set up mcaid_elig only
mcaid_claim_edges <- edges %>% 
  filter(str_detect(from, "mcaid_claim") | str_detect(to, "mcaid_claim"))

# See which nodes should be included
mcaid_claim_nodes <- node_filter(mcaid_claim_edges)
  
# Set up Graphviz text
mcaid_claim_gv <- gv_maker(node_df = mcaid_claim_nodes, edge_df = mcaid_claim_edges)

# Make graph
if (length(mcaid_claim_gv)) {
  grViz(mcaid_claim_gv)
} else {
  message("No tables in this category")
}
```

### Eligibility
```{r mcaid_elig, echo=FALSE}
# set up mcaid_elig only
mcaid_elig_edges <- edges %>% 
  filter(str_detect(from, "mcaid_elig") | str_detect(to, "mcaid_elig"))

# See which nodes should be included
mcaid_elig_nodes <- node_filter(mcaid_elig_edges)
  
# Set up Graphviz text
mcaid_elig_gv <- gv_maker(node_df = mcaid_elig_nodes, edge_df = mcaid_elig_edges)

# Make graph
if (length(mcaid_claim_gv)) {
  grViz(mcaid_claim_gv)
} else {
  message("No tables in this category")
}
```


## Medicare 
### Claims
```{r mcare_claim, echo=FALSE}
# set up mcare_elig only
mcare_claim_edges <- edges %>% 
  filter(str_detect(from, "mcare_claim") | str_detect(to, "mcare_claim"))

# See which nodes should be included
mcare_claim_nodes <- node_filter(mcare_claim_edges)
  
# Set up Graphviz text
mcare_claim_gv <- gv_maker(node_df = mcare_claim_nodes, edge_df = mcare_claim_edges)

# Make graph
if (length(mcare_claim_gv)) {
  grViz(mcare_claim_gv)
} else {
  message("No tables in this category")
}

```

### Eligibility
```{r mcare_elig, echo=FALSE}
# set up mcare_elig only
mcare_elig_edges <- edges %>% 
  filter(str_detect(from, c("\\.mcare_elig|mcare_mbsf")) | str_detect(to, c("\\.mcare_elig|mcare_mbsf")) ) %>% 
  filter(!(str_detect(from, c("mcaid_mcare")) | str_detect(to, c("mcaid_mcare"))) )

# See which nodes should be included
mcare_elig_nodes <- node_filter(mcare_elig_edges)
  
# Set up Graphviz text
mcare_elig_gv <- gv_maker(node_df = mcare_elig_nodes, edge_df = mcare_elig_edges)

# Make graph
if (length(mcare_elig_gv)) {
  grViz(mcare_elig_gv)
} else {
  message("No tables in this category")
}
```

## Identity linkage
```{r mcaid_mcare_pha_xwalk, echo=FALSE}
# set up mcaid_mcare_pha_xwalk only
mcaid_mcare_pha_xwalk_edges <- edges %>% 
  filter(str_detect(from, "xwalk_apde_mcaid_mcare_pha") | str_detect(to, "xwalk_apde_mcaid_mcare_pha")) %>% 
  filter(!(str_detect(from, c("final.xwalk_apde_mcaid_mcare_pha")) )) 

# See which nodes should be included
mcaid_mcare_pha_xwalk_nodes <- node_filter(mcaid_mcare_pha_xwalk_edges)
  
# Set up Graphviz text
mcaid_mcare_pha_xwalk_gv <- gv_maker(node_df = mcaid_mcare_pha_xwalk_nodes, edge_df = mcaid_mcare_pha_xwalk_edges)

# Make graph
if (length(mcaid_mcare_pha_xwalk_gv)) {
  grViz(mcaid_mcare_pha_xwalk_gv)
} else {
  message("No tables in this category")
}
```

## Medicaid/Medicare combined
### Claims
```{r mcaid_mcare_claim, echo=FALSE}
# set up mcaid_mcare_elig only
mcaid_mcare_claim_edges <- edges %>% 
  filter(str_detect(from, "mcaid_mcare_claim") | str_detect(to, "mcaid_mcare_claim"))

# See which nodes should be included
mcaid_mcare_claim_nodes <- node_filter(mcaid_mcare_claim_edges)
  
# Set up Graphviz text
mcaid_mcare_claim_gv <- gv_maker(node_df = mcaid_mcare_claim_nodes, edge_df = mcaid_mcare_claim_edges)

# Make graph
if (length(mcaid_mcare_claim_gv)) {
  grViz(mcaid_mcare_claim_gv)
} else {
  message("No tables in this category")
}
```

### Eligibility
```{r mcaid_mcare_elig, echo=FALSE}
# set up mcaid_mcare_elig only
mcaid_mcare_elig_edges <- edges %>% 
  filter(str_detect(from, "mcaid_mcare_elig") | str_detect(to, "mcaid_mcare_elig")) %>%
  filter(!(str_detect(from, c("final.mcaid_mcare_elig"))))

# See which nodes should be included
mcaid_mcare_elig_nodes <- node_filter(mcaid_mcare_elig_edges)
  
# Set up Graphviz text
mcaid_mcare_elig_gv <- gv_maker(node_df = mcaid_mcare_elig_nodes, edge_df = mcaid_mcare_elig_edges)

# Make graph
if (length(mcaid_mcare_elig_gv)) {
  grViz(mcaid_mcare_elig_gv)
} else {
  message("No tables in this category")
}
```


## Medicaid/Medicare/PH combined
### Claims
```{r mcaid_mcare_pha_claim, echo=FALSE}
# set up mcaid_mcare_pha_elig only
mcaid_mcare_pha_claim_edges <- edges %>% 
  filter(str_detect(from, "mcaid_mcare_pha_claim") | str_detect(to, "mcaid_mcare_pha_claim"))

# See which nodes should be included
mcaid_mcare_pha_claim_nodes <- node_filter(mcaid_mcare_pha_claim_edges)
  
# Set up Graphviz text
mcaid_mcare_pha_claim_gv <- gv_maker(node_df = mcaid_mcare_pha_claim_nodes, edge_df = mcaid_mcare_pha_claim_edges)

# Make graph
if (length(mcaid_mcare_pha_claim_gv)) {
  grViz(mcaid_mcare_pha_claim_gv)
} else {
  message("No tables in this category")
}
```

### Eligibility
```{r mcaid_mcare_pha_elig, echo=FALSE}
# set up mcaid_mcare_pha_elig only
mcaid_mcare_pha_elig_edges <- edges %>% 
  filter(str_detect(from, "mcaid_mcare_pha_elig") | str_detect(to, "mcaid_mcare_pha_elig"))

# See which nodes should be included
mcaid_mcare_pha_elig_nodes <- node_filter(mcaid_mcare_pha_elig_edges)
  
# Set up Graphviz text
mcaid_mcare_pha_elig_gv <- gv_maker(node_df = mcaid_mcare_pha_elig_nodes, edge_df = mcaid_mcare_pha_elig_edges)

# Make graph
if (length(mcaid_mcare_pha_elig_gv)) {
  grViz(mcaid_mcare_pha_elig_gv)
} else {
  message("No tables in this category")
}
```




